# QC and Genome Composition assessment


```{r, echo=FALSE, include=FALSE}
# load required libraries
up_packages = c("ggplot2", "grid", "gridExtra", "magrittr")
lapply(up_packages, require, character.only = TRUE)

# set working directory
mainDir = "/home/Andrea/Desktop/kveiks/manuscript/plotting_data/"
knitr::opts_knit$set(root.dir = mainDir)
```


## Reads QC and Filtering

Just trim low quality bases with Trimmomatic and discard short reads. I keep only properly paired reads and I discard the unpaired, since the latter are just a tiny fraction of my total data, they complicate the analysis and there is not much added value in them (at least for the analysis we run).

The file sample.lst just contains the names of my kveiks samples.


```{bash, eval=FALSE}
# generate FastQC report
for file in *.fastq; do
  ~/bin/FastQC/fastqc $file &
done

# LowQ base trimming and filtering
while read line; do
	java -jar ~/bin/trimmomatic/trimmomatic.jar PE \
		-threads 72 \
		-phred33 \
		$line.R1.fq.gz $line.R2.fq.gz \
		$line.R1.tr.fq.gz $line.R1.tr.un.fq.gz $line.R2.tr.fq.gz $line.R2.tr.un.fq.gz \
		SLIDINGWINDOW:10:30 \
		TRAILING:30 \
		MINLEN:50
done < sample.lst
```



## Genome Composition

We check the composition of the sequenced kveiks by aligning the reads to a set of Saccharomyces reference genomes.


```{r, echo=FALSE}
matrix.d = cbind(c("S. cerevisiae", "S. eubayanus", "S. kudriavzevi", "S. mikatae", "S. paradoxus", "S. uvarum"),
                 c("S288C_R64-2-1", "GCF_001298625.1", "gca_000167075", "SRX055454", "ASM207914v1", "ASM224264v1"),
                 c("Scere", "Seuba", "Skudr", "Smika", "Spara", "Suvar"))
colnames(matrix.d) = c("Species", "Genome Accession", "Label")

# print table
knitr::kable(
  matrix.d, caption = "Yeast Reference Genomes",
  booktabs = TRUE
)

```



First, we align the reads to the six _Saccharomyces_ reference genomes. Then we create 10kb windows with ```bedtools``` and we calculate the average coverage for each window.


```{bash, eval=FALSE}
## index the reference genome
~/bin/bwa/bwa index Saccharomyces_RefGen.fa

## reads alignment
while read line ; do
	bwa mem -t 72 \
		-K 100000000 \
		Saccharomyces_RefGen.fa \
	  "${line}".R1.tr.fq.gz "${line}".R2.tr.fq.gz \
    > "${line}".align.sam
  samtools view -@ 72 -Sb \
    "${line}".align.sam > "${line}".align.bam
  samtools sort -@ 72 
    "${line}".align.bam "${line}".align.sort
done < ../sample.lst 

## bam2bed
# prepare reference genomes
samtools-1.9/samtools faidx Saccharomyces_RefGen.fa
gatk CreateSequenceDictionary --R Saccharomyces_RefGen.fa
cut -f 2,3 Saccharomyces_RefGen.dict \
  | tail -n +2 \
  | sed 's/LN://g' \
  | sed 's/SN://g' \
  > Saccharomyces_RefGen.bedchor
bedtools makewindows -g Saccharomyces_RefGen.bedchor \
	-w 1000 > Saccharomyces_RefGen.1kb_win.tab

# calculate mean coverage
while read line; do
	bedtools coverage -a Saccharomyces_RefGen.1kb_win.tab \
	-b "${line}".align.sort.bam -mean > "${line}".align.sort.1kb_cov.bed;
done

# format output
for file in *.1kb_cov.bed ; do
	sed -i 's/Scere_/Scere\t/g' $file;
	sed -i 's/Seuba_/Seuba\t/g' $file;
	sed -i 's/Skudr_/Skudr\t/g' $file;
	sed -i 's/Smika_/Smika\t/g' $file;
	sed -i 's/Spara_/Spara\t/g' $file;
	sed -i 's/Suvar_/Suvar\t/g' $file;
	python bed_remove0coverage.py --input $file |\
    	sed '/^[[:space:]]*$/d' > $(basename $file .bed).no0.bed;
done

for file in *.no0.bed; do
	NAME=$(basename $file .align.sort.1kb_cov.no0.bed);
	sed -i "s/^/$NAME\t/" $file;
done

cat *.1kb_cov.no0.bed > Viking.species.cov.bed


# generate per chromosomes coverages
while read line; do
	OLD=$(echo $line | cut -d " " -f 1)
	NEW=$(echo $line | cut -d " " -f 2)
	sed -i "s/$OLD\t/$NEW\t/g" Viking.species.cov.chr.bed
done < ref_chr_names.lst 

# clean output
for file in *.no0.bed; do
	cp $file $file.1;
	while read line; do
		OLD=$(echo $line | cut -d " " -f 1);
		NEW=$(echo $line | cut -d " " -f 2);
		sed -i "s/$OLD\t/$NEW\t/g" $file.1;
	done < ../00_ref_genomes/ref_chr_names.lst;
	grep Chr $file.1 > $(basename $file .bed).chr.bed;
	rm $file.1 &
done
```


We then remove windows with 0 coverage to have a clearer figure, and we plot the coverages with transforming the y axis in log10 scale, since Krogerus kveiks have much higher coverage than the strains we sequenced.



```{r, fig.cap="Coverage of Reference Genomes", fig.height=9,  warning=FALSE, message=FALSE}
data_cov = read.delim(paste(mainDir, "Viking.species.cov.chr.bed", sep = ""), header = FALSE)
data_cov$V7 = ifelse(data_cov$V6 >= 1, "Up", "Down")

# ylim = 1500
ggplot(data_cov, aes(x = V2, y = V6)) +
  geom_jitter(aes(color = V7),
              show.legend = FALSE,
              alpha = 0.035,
              shape = ".",
              position = position_jitter(0.475)) +
  geom_hline(yintercept = 1, linetype = 2, color = "red") +
  coord_cartesian(expand = FALSE) +
  scale_color_manual(values  = c("grey75", "cyan4")) + 
  scale_y_continuous(limits = c(-4, 9000),
                     trans = "log2",
                     breaks = c(1, 10, 100, 1000),
                     labels = scales::comma) +
  labs(title = "Coverage of Reference Genomes",
       x = "Reference Species",
       y = "Coverage") +
  facet_wrap(~V1, scales = "free_y") +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        panel.background = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "gray75", size = 0.5),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.65))
```



All sequenced farmhouse yeasts are pure _S. cerevisiae_ strains, except for Muri and k7R25 that are _S. cerevisiae_, _S. eubayanus_ and _S. uvarum_ triple hybrids.

We can dig in into the coverage of the single chromosomes and visualize the extent of hybridization and eventual aneuploidies. 



```{r, fig.cap="Sub-genomes coverages", warning=FALSE, message=FALSE}

### PER SAMPLE COVERAGE
sample_list = unique(data_cov$V1)
r = list()

# I iterate through the samples
for (i in 1:length(sample_list)){
  
  # I read the sample name and restrict the plot to that sample
  name = sample_list[i]
  file = paste(mainDir, "/01_start_bed/",name,".align.sort.1kb_cov.no0.chr.bed", sep = "")
  dataset = read.delim(file, header = FALSE)
  
  # plot
  p = ggplot(dataset, aes(V5, V6, colour = V2)) +
    geom_point(size = 0.2) +
    scale_y_continuous(limits = c(1.5, 9000),
                       trans = "log2",
                       breaks = c(1, 10, 100, 1000),
                       labels = scales::comma) +
    scale_x_discrete(breaks = (seq(0,2e6,10000))) +
    coord_cartesian(expand = FALSE) +
    facet_grid(V2~dataset$V3,
               scales = "free",
               space = "free_x",
               switch = "y") +
    labs(title = name,
         x = "Chromosome",
         y = "Genome coverage",
         color = "Genomes") +
    theme(plot.title = element_text(hjust = 0.5, size = 20),
          legend.position = "none",
          panel.background = element_blank(),
          panel.spacing.x = unit(0, "lines"),
          panel.border = element_rect(colour = "black", size = 1, fill = NA),
          strip.background = element_rect(colour = "black", size = 1, fill = NA))
  r[[name]] = p
  
  if(name %in% c("14R30", "Muri", "k7R25")){
    print(p)
  }
}


```


Muri and k7R25 are triple hybrids _S. cerevisiae_ X _S. eubayanus_ X _S. uvarum_. What is extremely interesting is that they share the same genomic composition in terms of origin of the chromosomes, but they have distinct CNVs patterns. I would suggest a common origin for such hybrid.




## Ploydies estimation

We can estimates Kveiks ploidy by looking at the k-mer frequency distributions of the sequenced reads obtained from whole genome sequencing.

```{bash, eval=FALSE}
# calculate kmer freq
for file in 01_alignments/*bam; do
  ~/bin/ntCard/bin/ntcard --threads=64 --kmer=41 \
    --pref=$(basename $file .align.sort.md.r.bam).kmer41freq $file;
done
# merge outputs
for file in *.hist; do
  tail -n +3 $file > $file.1;
  mv $file.1 $file;
  cat $file | sed "s/^/$(basename $file .kmer41freq_k41.hist)\t/g";
done > Viking.species.kmer41freq.txt
```



Kmer frequency distribution plots can suggest us the degree of heterozygosity of kveiks genomes. All Kveiks seems to have a high degree of heterozygosity, except for hybrid strains Muri and k7R25. This support the idea that the sequenced kveiks so far are tetraploid _S. cerevisiae_, while Muri and k7R25 are triploid hybrids _S. cerevisiae_ X _Seubayanus_ X _S. uvarum_.


```{r, fig.cap="K-mer frequencies distribution", fig.height=9, warning=FALSE, message=FALSE}

data_k = read.delim(paste(mainDir, "Viking.species.kmer41freq.mod.txt", sep = ""), header = FALSE)
data_k$V2 = as.numeric(data_k$V2)
data_k$V3 = as.numeric(data_k$V3)

# plot kmer distributions
ggplot(data_k, aes(x = V2, y = V3)) +
  geom_line(color = "steelblue", size = 1) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  facet_wrap(~V1, scales = "free") +
  labs(title = "K-mer frequencies distribution",
       x = "k-mer depth",
       y = "k-mer frequency") +
  theme(plot.title = element_text(size = 20, hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8))
```



We compare the predicted ploidy results with the ones of 18 _S. cerevisiae_ industrial strains that will be used across this whole study as comparison.


```{r, add_yeasts, echo=FALSE}
matrix.d = cbind(c("X1002", "X1014", "X1018", "X1030", "X1129", "X5508",
                   "X1005", "X1007", "X1118", "X1131", "X1148", "X7078-B",
                   "X1046", "X4009", "X4037", "X4040", "X4042", "X4051"),
                 c("Beer1", "Beer1", "Beer1", "Beer1", "Beer1", "Beer1",
                   "Beer2", "Beer2", "Beer2", "Beer2", "Beer2", "Beer2",
                   "Wine", "Wine", "Wine", "Wine", "Wine", "Wine"))
colnames(matrix.d) = c("Strain", "Industrial niche")

# print table
DT::datatable(matrix.d, 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "Yeast Reference Genomes", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))

```



```{r, fig.cap="K-mer frequencies distribution \n Industrail strains", fig.height=6, warning=FALSE, message=FALSE}

data_ay = read.delim(paste(mainDir, "additional_yeasts.kmer41freq.txt", sep = ""), header = FALSE)
data_ay$V2 = as.numeric(data_ay$V2)
data_ay$V3 = as.numeric(data_ay$V3)
data_ay$V4 = ifelse(data_ay$V1 == "X1002" | data_ay$V1 == "X1014" | data_ay$V1 == "X1018" | data_ay$V1 == "X1030" | data_ay$V1 == "X1129" | data_ay$V1 == "X5508", "Beer1",
                    ifelse(data_ay$V1 == "X1005" | data_ay$V1 == "X1007" | data_ay$V1 == "X1118" | data_ay$V1 == "X1131" | data_ay$V1 == "X1148" | data_ay$V1 =="x7078-B", "Beer2", "Wine"))
data_ay$V1 = factor(data_ay$V1,
                    levels = c("X1002", "X1014", "X1018", "X1030", "X1129", "X5508",
                               "X1005", "X1007", "X1118", "X1131", "X1148", "x7078-B",
                               "X1046", "X4009", "X4037", "X4040", "X4042", "X4051"))

# plot kmer distributions
ggplot(data_ay, aes(x = V2, y = V3)) +
  geom_line(aes(color = V4), size = 1) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  xlim(10, 200) +
  ylim(0, 515000) +
  facet_wrap(~V1, scales = "free", ncol = 6) +
  labs(title = "K-mer frequencies distribution",
       colour = "Industrial\n niche",
       x = "k-mer depth",
       y = "k-mer frequency") +
  theme(plot.title = element_text(size = 20, hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8))
```





```{r, fig.cap="Chromosome coverages in Industrail strains", fig.height=6, warning=FALSE, message=FALSE}

data_ax = read.delim(paste(mainDir, "additional_yeasts.chr_coverage.txt", sep = ""), header = FALSE)
data_ax$V3 = as.numeric(data_ax$V3)
data_ax$V4 = as.numeric(data_ax$V4)
data_ax$V5 = as.numeric(data_ax$V5)

data_ax$V2 = factor(data_ax$V2,
                    levels = c("I", "II", "III", "IV", "V", "VI", "VII", "VIII",
                               "IX", "X", "XI", "XII", "XIII", "XIV", "XV", "XVI"))

data_ax$V6 = ifelse(data_ax$V1 == "X1002" | data_ax$V1 == "X1014" | data_ax$V1 == "X1018" | data_ax$V1 == "X1030" | data_ax$V1 == "X1129" | data_ax$V1 == "X5508", "Beer1",
                    ifelse(data_ax$V1 == "X1005" | data_ax$V1 == "X1007" | data_ax$V1 == "X1118" | data_ax$V1 == "X1131" | data_ax$V1 == "X1148" | data_ax$V1 == "X7078-B", "Beer2", "Wine"))
data_ax$V1 = factor(data_ax$V1,
                    levels = c("X1002", "X1014", "X1018", "X1030", "X1129", "X5508",
                               "X1005", "X1007", "X1118", "X1131", "X1148", "X7078-B",
                               "X1046", "X4009", "X4037", "X4040", "X4042", "X4051"))

# plot Beer1 coverages
ggplot(subset(data_ax, V6 == "Beer1"), aes(x = V3, y = V5)) +
  geom_point(color = "#F8766D", size = 0.2) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  #xlim(10, 200) +
  #ylim(0, 515000) +
  scale_y_continuous(limits = c(0, 1300)) +
  scale_x_discrete(breaks = (seq(0,2e6,10000))) +
  coord_cartesian(expand = FALSE) +
  facet_grid(V2~V3,
             scales = "free",
             space = "free_x",
             switch = "y") +
  labs(title = "Chromosome coverage - Beer1",
       x = "Chromosome",
       y = "Coverage") +
  theme(plot.title = element_text(size = 20, hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8))
```


