# Horizontal Gene Transfer

```{r, echo=FALSE, include=FALSE}
# load required libraries
up_packages = c("circlize", "clusterProfiler", "ComplexHeatmap", "ggplot2", "grid", "gridExtra", "magrittr", "RColorBrewer",
                "org.Sc.sgd.db")
lapply(up_packages, require, character.only = TRUE)

# set working directory
mainDir = "/home/andrea/19ADC99/vikingsproject"
knitr::opts_knit$set(root.dir = mainDir)

# retrieve S. cerevisiae genome annotation
Scere_DB = biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
                            dataset="scerevisiae_gene_ensembl")
Scere_DB_table = biomaRt::getBM(attributes = c("ensembl_gene_id", "ensembl_peptide_id", "external_gene_name", "entrezgene_id", "description", "chromosome_name", "start_position", "end_position"),
                                mart = Scere_DB)

# build list of reference databases used for annotation steps
ref_DB_list = c("org.Sc.sgd.db", "yeast", "scerevisiae")

# create yeasts GO terms universe
GO_universe = data.frame(matrix(nrow = nrow(Scere_DB_table), ncol = 2))
names(GO_universe) = c("ENSEMBL", "EntrezID")
GO_universe$ENSEMBL = Scere_DB_table$ensembl_gene_id
# populate GO universe
for(k in 1:nrow(GO_universe)){
  # force to get only the first term if multiple are retrieved (sic!)
  ENSEMBL = GO_universe[k, 1]
  GO_universe[k, 2] = tryCatch(
    Scere_DB_table[which(Scere_DB_table[, 1] == ENSEMBL),  c(4)][[1]],
    error = function(e) { NA }
  )
}
GO_universe = as.character(c(GO_universe[!is.na(GO_universe$EntrezID), ]$EntrezID))
```


To identify genuine Bacterial (or Fungal) Horizontal Gene Transfer, we apply the following protocol:

- sequence similarity search against non-redundant proteins database, including taxonomic annotations
- identify protein coding genes with best hits (top 5) to Bacterial proteins
- select assembled contigs where putative bacterial genes are
- discard short contigs with only one bacterial hit (Noise)
- if putative bacterial gene flanked by Eukaryotic genes, manual sequence similarity search to confirm a _bona fide_ hit

To have internal controls in this (and in the following analyses), we will add the ones of 18 _S. cerevisiae_ industrial strains as well.


## Taxonomic annotation of predicted genes

For each predicted protein-coding gene, we can make a sequence similarity search against the non-redundant protein database at NCBI (nr). To assign the taxonomy id of the nr BLAST hit, we need the ```prot.accession2taxid.gz``` and ```nodes.dmp``` files provided by NCBI taxonomy. This can help us identify horizontal gene transfer and traces of contamination in the library prep of yeast samples.

For each sample, first we do a blast search using DIAMOND, and then we associate to the protein_id of the top 5 best hits to the corresponding taxonomy the python library ete and customs scripts. BLAST is run against nr, excluding _Saccharomyces cerevisiae_, so that we can judge if the protein was present in other yeasts, or if it is indeed of bacterial origin.


```{bash, eval=FALSE}
# BLAST search
for file in *.all.maker.proteins.fasta; do
  ~/bin/DIAMOND/diamond blastp \
    --query $file \
    --db ~/taxonomy/nr \
    --taxonmap ~/taxonomy/prot.accession2taxid.gz \
    --taxonnodes ~/taxonomy/nodes.dmp \
    --threads 70 \
    --sensitive \
    --max-target-seqs 5 \
    --outfmt 6 qseqid sseqid pident length mismatch gapopen qstart qend
    sstart send evalue bitscore staxids > $file.diamond
done

# BLAST to taxid
for file in *.diamond; do
  ~/anaconda_ete/bin/python3.6 Vikings.tax.topath.py --input $file > $file.tax
done
for file in *.tax ; do
  while read line ; do
    grep $line $file | cut -f 1,5;
    echo;
  done < <(grep Bacteria $file | cut -f 1 | sort -u) > $file.bact
done
```


Now we have the full taxonomic annotation of the five best hits for each of the protein coding genes we annotated on the kveiks.



## Evidence for bacterial HGT


```{bash, eval=FALSE}
# filter annot.gff files
for file in *.gff ; do
  grep CDS $file > $(basename $file .gff).CDSonly.gff;
  gzip $file;
done

# identify putative bacterial genes and the assembled contigs harbouring them
for file in *.tax; do cut -f 1,5 $file > $file.all; done
for file in *.all; do grep Bact $file > $(basename $file .all).bact; done
for file in *.bact; do
  while read line; do
    if [[ $line = "" ]]; then
      continue;
    else
      GENE=$(echo $line | cut -d " " -f 1 );
      CHR=$(grep "${GENE}" $(basename $file .maker.proteins.fasta.diamond.tax.bact).CDSonly.gff | cut -f 1 );
      grep "${CHR}" $(basename $file .bact).all \
        | uniq \
        | grep -C5 "${GENE}" \
        | grep -v "${GENE}" \
        >> $(basename $file .maker.proteins.fasta.diamond.tax.bact)."${GENE}".tmp;
    fi;
  done < <(cut -f 1 $file | sort -u);
done

# pull the putative bacterial contigs
for file in *.bact; do
  cut -f 1 $file | sort -u > $file.lst;
  perl ~/scripts/SelectList_Fasta.pl \
    $(basename $file .diamond.tax.bact) \
    $file.lst \
    > $file.fa;
done

# check if putative bacterial transcripts are flanked by Euk genes
python3.6 Vikings.check_eukbact_contigs.1.py --input ../samples.lst
for file in *HGT.table; do uniq $file > temp; mv temp $file ; done
for file in *HGT.table; do
  python3 ../Vikings.Bactmatch.py --input $file | cut -f 2 > temp;
  cp temp temp2;
  paste temp temp2 | sed 's/\t/\|\>/g' > $file.lst;
  rm temp temp2;
done
for file in *HGT.table.lst; do
  perl ~/scripts/SelectList_Fasta.pl \
    $(basename $file .HGT.table.lst).bact.fa \
    $file \
    > $(basename $file .lst).fa;
done
```

**THERE ARE NO EVIDENCES FOR SIGNIFICANT BACTERIAL HORIZONTAL GENE TRANSFER.**


## Evidence for HGT from Ascomycota

We can check the presence of fungal non-Saccharomyces genes and operons starting from the DIAMOND blast results we have.


```{bash, eval=FALSE}
# Select non Saccharomyces genes
# filter for genes with no top hits to Saccharomyces
for file in *tax; do
  grep Ascomycota $file | grep -v Saccharomyces > $file.Asco;
  while read line; do
    grep $line $file > $file.Asco.all;
    if [[ $(grep $line $file | grep Saccharomyces) ]]; then
      continue;
    else
      echo $line;
    fi;
  done < <(cut -f 1 $file.Asco | sort -u) > $file.Asco.candidates;
done

# retrieve fasta sequences
for file in *candidates; do
perl ~/scripts/SelectList_Fasta.pl \
  ../../11_domestication/00_prot_DB/$(basename $file
  .diamond.tax.Asco.candidates) \
  $file > $file.fa;
done

# generate table stats
for i in 2 3; do
  wc -l *candidates \
    | sed "s/.SPAdes.*//g" \
    | sed "s/.aa.*//g" \
    | sed "s/.contigs.*//g" \
    | grep -v total \
    | sed "s/  / /g" \
    | sed "s/  / /g" \
    | cut -f $i -d " " \
    > temp.$i;
done;

paste temp.3 temp.2 > Vikings.Asco.HGT.stats;

# clean up
rm temp.2 temp.3
```



```{r, echo=FALSE}
matrix.d = cbind(c("14R30", "14R6", "16R23", "16R37", "17P5", "19R18", "1R16", "21P1", "21R38",
                   "27R17", "28P1", "28P6", "28R21", "28R33", "28R8", "2R23", "38R16", "39R20", 
                   "3R11", "40R14", "40R1", "40R20", "41R10", "41R15", "42R20", "42R31", "44R32", "44R7", 
                   "45P5", "45R11", "46R12", "46R37", "6R15", "7R7", "8R19", "9R40", "Granvin1", 
                   "Hornindal1", "Hornindal2", "k7R25", "Laerdal2", "Muri", "SortdalEbbe1", "Voss1",
                   "X1002", "X1014", "X1018", "X1030", "X1129", "X5508",
                   "X1005", "X1007", "X1118", "X1131", "X1148", "X7078-B",
                   "X1046", "X4009", "X4037", "X4040", "X4042", "X4051"),
                 c("76", "74", "4", "22", "92", "90", "92", "106", "4", "73", "4", "6", "4", "4", "5",
                   "90", "56", "8", "4", "5", "4", "96", "4", "6", "5", "13", "90", "76", "14", "13",
                   "16", "11", "109", "81", "57", "58", "4", "5", "4", "21", "5", "15", "6", "5",
                   "3", "3", "NA", "4", "4", "3", "NA", "3", "3", "NA", "3", "NA", "6", "NA", "3", "NA", "NA", "4"),
                 c("Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse",
                   "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", 
                   "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", 
                   "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", 
                   "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse", "Farmhouse",
                   "Beer1", "Beer1", "Beer1", "Beer1", "Beer1", "Beer1",
                   "Beer2", "Beer2", "Beer2", "Beer2", "Beer2", "Beer2", 
                   "Wine", "Wine", "Wine", "Wine", "Wine", "Wine"),
                 c("North-West Norway", "North-West Norway", "North-West Norway", "North-West Norway", "North-West Norway",
                   "North-West Norway", "North-West Norway", "North-West Norway", "North-West Norway", "North-West Norway",
                   "North-West Norway", "North-West Norway", "North-West Norway", "North-West Norway", "South-West Norway",
                   "South-West Norway", "South-West Norway", "South-West Norway", "South-West Norway", "South-West Norway",
                   "South-West Norway", "South-West Norway", "South-West Norway", "South-West Norway", "Central-Eastern Norway",
                   "Central-Eastern Norway", "Central-Eastern Norway", "Central-Eastern Norway", "Central-Eastern Norway",
                   "Central-Eastern Norway", "Latvia", "Latvia", "Latvia", "Latvia", "Lithuania", "Lithuania", "Lithuania",
                   "Lithuania", "Russia", "Russia", "Russia", "Russia", "Other", "Other",
                   "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA"))
colnames(matrix.d) = c("Strains", "Putative Ascomicota HGT", "Industrial niche", "Geographical origin")

# print table
DT::datatable(matrix.d, 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "Farmhouse yeasts Ascomycota HGT genes", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))
```



Kveik strains seems to have a significant higher number of Ascomycetes genes than the other industrial strains analysed. Are there additional proteins shared between all the kveiks (or by some kveiks with same geographical origin?). We can group these proteins in gene families and see if we see a common pattern. If not, we can BLAST back the proteins and see which specific genes have been acquired.


```{bash, eval=FALSE}
# group candidates in gene families
~/bin/OrthoFinder-2.3.3/orthofinder -f ./ \
  -t 20 \
  -M msa \
  -S diamond \
  -A mafft \
  -T iqtree \
  -I 2
```



From OrthoFinder output we can identify 49 orthogroups containing four genes or more. We visualize them as heatmap, anc we can see that kveiks strains have a higher number of common additional orthogroups that are absent from S288C and industrial strains. Interestingly, the distribution pattern does not seem to overlap with the geographical isolation of the kveik culture.



```{r, fig.cap="Ascomycota HTG orthogroup candidates", fig.height=10, warning=FALSE, message=FALSE}
# import table
heatfile = read.delim(paste(mainDir, "Vikings.Asco.HGT.heatmap.tab", sep = ""), header = FALSE)
heatfile$V2 = stringr::str_replace_all(heatfile$V2, "x", "X")
heatfile = heatfile[which(heatfile$V2 %in% c("X1002", "X1014", "X1030", "X1129", "X5508", "X1007", "X1118", "X1148", "X1046", "X4037",
                                "X4051", "Voss1", "SortdalEbbe1", "Muri", "Laerdal2", "k7R25", "Hornindal2", "Hornindal1",
                                "Granvin1", "9R40", "8R19", "7R7", "6R15", "46R37", "46R12", "45R11", "45P5", "44R7", "44R32",
                                "42R31", "42R20", "41R15", "41R10", "40R20", "40R1", "40R14", "3R11", "39R20", "38R16", "2R23",
                                "28R8", "28R33", "28R21", "28P6", "28P1", "27R17", "21R38", "21P1", "1R16", "19R18", "17P5",
                                "16R37", "16R23", "14R6", "14R30")), ]

# relevel
heatfile$V2 = factor(heatfile$V2,
                     levels = c("41R10", "21R38", "9R40", "17P5", "SortdalEbbe1", "3R11", "21P1", "41R15", "Hornindal1",
                                "Hornindal2", "1R16", "2R23", "8R19", "Muri",
                                "k7R25", "38R16", "44R32", "19R18", "44R7", "6R15", "Laerdal2", "7R7", "14R6", "14R30",
                                "27R17", "28P1", "28P6", "28R21", "28R33", "28R8",
                                "42R20", "42R31", "45P5", "45R11", 
                                "46R12", "46R37", "16R23", "16R37",
                                "39R20", "40R14", "40R1", "40R20",
                                "Granvin1", "Voss1",
                                "X1002", "X1014", "X1030", "X1129", "X5508", "X1007", "X1118", "X1148", "X1046", "X4037",
                                "X4051"))

# set color labels
col_label = fills = c("#0571B0", "#0571B0", "#0571B0", "#0571B0", "#0571B0", "#0571B0", "#0571B0", "#0571B0",
                      "#0571B0", "#0571B0", "#0571B0", "#0571B0", "#0571B0", "#92C5DE", "#92C5DE",
                      "#92C5DE", "#92C5DE", "#92C5DE", "#92C5DE", "#92C5DE", "#92C5DE", "#92C5DE", "#92C5DE",
                      "#008470", "#008470", "#008470", "#008470", "#008470", "#008470", "#FFDA00", "#FFDA00",
                      "#FFDA00", "#FFDA00", "#FBA01D", "#FBA01D", "#FBA01D", "#FBA01D", "#A6611A", "#A6611A",
                      "#A6611A", "#A6611A", "grey75", "grey75", "#51CD34", "#51CD34", "#51CD34", "#51CD34",
                      "#51CD34", "#51CD34", "#2E97E6", "#2E97E6", "#2E97E6", "#2E97E6", "#2E97E6", "#2E97E6",
                      "#F2BA00", "#F2BA00", "#F2BA00", "#F2BA00", "#F2BA00", "#F2BA00", "#F2BA00", "#F2BA00")


# prepare heatmap
ggplot(heatfile) +
  geom_tile(aes(x = V1, y = V2, fill = V3), color = "grey75") +
  scale_fill_gradientn(na.value = "white", limits = c(0, 2),
                       colours = c("white", "steelblue", "steelblue"),
                       breaks = c(0, 1, 2)) +
  coord_cartesian(expand = FALSE) +
  labs(fill = "Gene presence",
       y = "Strain",
       x = "Orthologous groups") +
  theme(axis.title = element_text(size = 16),
        axis.text.x = element_text(angle = 90, hjust = 0.95, vjust = 0.5),
        axis.text.y = element_text(size = 16, colour = col_label),
        legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, size = 0.75)) +
  annotate(xmin = 25.5, xmax = 32.5, 
           ymin = -Inf, ymax = Inf, 
           geom = "rect", alpha = 0,
           colour = "red", linewidth = 1.5)
```



Very well. Now, what are the gene families that are transferred to kveiks strains? Are they genuine HGT, or it is just artifacts from heuristic sequence similarity search? Let's do manual BLAST to NCBI for the Orthogroups and see what pops up.

```{r, echo=FALSE}
# read best hits table
matrix.d = read.delim(paste(mainDir, "Vikings.Asco.HGT.besthits.tab", sep = ""), header = FALSE)
colnames(matrix.d) = c("Orthogroups", "# strains", "# kveiks", "Best hit to", "Accession", "Annotation")

# print table
DT::datatable(matrix.d, 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "Inferred Ascomycota HGT taxonomic origin", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))
```



We can clearly identify two operons (OG0000024-OG0000031) that were transferred from _Zygosaccharomzces parabailli_ to kveik strains 16R37, 42R31, 45P5, 45R11 and 46R37.

These strains come all from a small geographic area (Latvia [42R31, 45P5, 45R11] and Lithuania [16R37, 46R37]), suggesting a common origin of the HGT event that then spread. Interestingly, other isolates from the same culture (i.e.: 16R23, 42R20, 46R12) do not present such a HGT, supporting the idea of heterogeneous kvieks cultures.


### _Zygosaccharomyces parabailli_ transferred operon

The transferred operons code for the genes: YNR058W, YHR204W, YHR205W, YHR206W, YHR207C, YHR208W, YER182W, YER183C.


```{r, echo=FALSE}
# annotate HGT genes
HGT_genes = data.frame("ENSEMBL" = c("YNR058W", "YHR204W", "YHR205W", "YHR206W", "YHR207C", "YHR208W", "YER182W", "YER183C"),
                       "ENTREZ" = rep(NA, 8), "SYMBOL" = rep(NA, 8), "Description" = rep(NA, 8))
for(k in 1:nrow(HGT_genes)){
  ENSEMBL = HGT_genes$ENSEMBL[[k]]
  ENTREZ = Scere_DB_table[which(Scere_DB_table$ensembl_gene_id == ENSEMBL), ]$entrezgene_id
  SYMBOL = Scere_DB_table[which(Scere_DB_table$ensembl_gene_id == ENSEMBL), ]$external_gene_name
  Description = Scere_DB_table[which(Scere_DB_table$ensembl_gene_id == ENSEMBL), ]$description
  HGT_genes[k, "ENTREZ"] = ENTREZ
  HGT_genes[k, "SYMBOL"] = SYMBOL
  HGT_genes[k, "Description"] = Description
}

# print table
DT::datatable(HGT_genes, 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "HGT Zygosaccharomzces parabailli operon", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))

```


#### Overrepresented GO BP terms

```{r, warning=FALSE, message=FALSE}
### run over-represented analysis
enriched_GOs = enrichGO(gene = HGT_genes$ENTREZ,
                               universe = GO_universe,
                               OrgDb = ref_DB_list[[1]],
                               ont = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff = 0.05,
                               qvalueCutoff = 0.05)

# print table
DT::datatable(enriched_GOs@result[which(enriched_GOs@result$p.adjust <= 0.05), ], 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "Overrepresented BP terms in Zygosaccharomzces parabailli transferred operon", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))
```


```{r, fig.cap="Dotplot of overrepresented BP terms in Zygosaccharomzces parabailli transferred operon", warning=FALSE, message=FALSE}
# plot GO dotplot
p0 = enrichplot::dotplot(enriched_GOs, showCategory = nrow(enriched_GOs@result)) +
  ggtitle("Zygosaccharomzces parabailli transferred operon over-represented GO BP terms sets") +
  scale_colour_gradientn(colours = rev(colorRampPalette(brewer.pal(7, "Reds"))(37))) +
  guides(colour = guide_colorbar(reverse = TRUE)) +
  theme(plot.title = element_text(size = 22, hjust = 0.5),
        legend.position = "bottom",
        strip.background = element_rect(fill = "grey85", colour = "black"),
        strip.text = element_text(size = 12))
print(p0)
```



```{r, fig.cap="Network of overrepresented BP terms in Zygosaccharomzces parabailli transferred operon", warning=FALSE, message=FALSE}
# plot GO network
p1 = goplot(enriched_GOs) +
            scale_colour_gradientn(colours = rev(colorRampPalette(brewer.pal(7, "Reds"))(37))) +
            labs(title = "Zygosaccharomzces parabailli transferred operon over-represented GO BP terms sets") +
            theme(plot.title = element_text(size = 20, hjust = 0.5))
print(p1)
```


#### Overrepresented GO CC terms

```{r, warning=FALSE, message=FALSE}
### run over-represented analysis
enriched_GOs = enrichGO(gene = HGT_genes$ENTREZ,
                               universe = GO_universe,
                               OrgDb = ref_DB_list[[1]],
                               ont = "CC",
                               pAdjustMethod = "BH",
                               pvalueCutoff = 0.05,
                               qvalueCutoff = 0.05)
```


#### Overrepresented GO MF terms

```{r, warning=FALSE, message=FALSE}
### run over-represented analysis
enriched_GOs = enrichGO(gene = HGT_genes$ENTREZ,
                               universe = GO_universe,
                               OrgDb = ref_DB_list[[1]],
                               ont = "MF",
                               pAdjustMethod = "BH",
                               pvalueCutoff = 0.05,
                               qvalueCutoff = 0.05)

# print table
DT::datatable(enriched_GOs@result[which(enriched_GOs@result$p.adjust <= 0.05), ], 
              extensions = c("FixedColumns", "FixedHeader"),
              caption = "Overrepresented MF terms in Zygosaccharomzces parabailli transferred operon", 
              options = list(scrollX = TRUE, paging = TRUE, fixedHeader = FALSE))
```


```{r, fig.cap="Dotplot of overrepresented MF terms in Zygosaccharomzces parabailli transferred operon", warning=FALSE, message=FALSE}
# plot GO dotplot
p0 = enrichplot::dotplot(enriched_GOs, showCategory = nrow(enriched_GOs@result)) +
  ggtitle("Zygosaccharomzces parabailli transferred operon over-represented GO MF terms sets") +
  scale_colour_gradientn(colours = rev(colorRampPalette(brewer.pal(7, "Reds"))(37))) +
  guides(colour = guide_colorbar(reverse = TRUE)) +
  theme(plot.title = element_text(size = 22, hjust = 0.5),
        legend.position = "bottom",
        strip.background = element_rect(fill = "grey85", colour = "black"),
        strip.text = element_text(size = 12))
print(p0)
```



```{r, fig.cap="Network of overrepresented MF terms in Zygosaccharomzces parabailli transferred operon", warning=FALSE, message=FALSE}
# plot GO network
p1 = goplot(enriched_GOs) +
            scale_colour_gradientn(colours = rev(colorRampPalette(brewer.pal(7, "Reds"))(37))) +
            labs(title = "Zygosaccharomzces parabailli transferred operon over-represented GO MF terms sets") +
            theme(plot.title = element_text(size = 20, hjust = 0.5))
print(p1)
```



#### Overrepresented KEGG Pathways


```{r, warning=FALSE, message=FALSE}
# over-represented KEGG
enriched_KEGG = enrichKEGG(gene = HGT_genes$ENTREZ,
                           keyType = "ncbi-geneid",
                           organism = "sce",
                           pvalueCutoff = 0.05)
```



#### Overrepresented Reactome Pathways


```{r, warning=FALSE, message=FALSE}
# over-represented PA
enrichedPA = ReactomePA::enrichPathway(gene = HGT_genes$ENTREZ,
                                       pvalueCutoff = 0.05,
                                       organism = ref_DB_list[[2]])
```





