# High-throughput phenotyping

```{css, echo=FALSE}

pre {
  max-height: 300px;
}

div.sourceCode {
  overflow: scroll;
}

.bash-code {
  font-style: italic;
}


```


```{r, echo=FALSE, include=FALSE}
# load required libraries
up_packages = c("ape", "aplot", "dplyr", "ggnewscale", "ggplot2", "ggpubr", "ggtree", "ggtreeExtra",
                "grid", "gridExtra", "gtable", "naturalsort", "phytools", "plyr", "RColorBrewer",
                "reshape", "reshape2", "stringr", "treeio")
lapply(up_packages, require, character.only = TRUE)

# set working directory
mainDir = "/home/andrea/19ADC99/vikingsproject"
knitr::opts_knit$set(root.dir = mainDir)

```


## On this page

Biological insights and take-home messages are at the bottom of the page at section Lesson Learnt: @sec-lesson-learnt-feature-selection.

* Here


## High-throughput phenotyping of farmhouse and industrial yeasts



## Fermentations





## Lessons Learnt {#sec-lesson-learnt-feature-selection}

Based on the  we have learnt:

* Fr


## Session Information

:::{.callout-note collapse=true}

```{r, echo=FALSE}
utils::sessionInfo()
```

:::