# QC and Genome Composition assessment

```{css, echo=FALSE}

pre {
  max-height: 300px;
}

div.sourceCode {
  overflow: scroll;
}

.bash-code {
  font-style: italic;
}


```


```{r, echo=FALSE, include=FALSE}
# load required libraries
up_packages = c("ggplot2", "grid", "gridExtra", "magrittr")
lapply(up_packages, require, character.only = TRUE)

# set working directory
mainDir = "/home/andrea/19ADC99/vikingsproject"
knitr::opts_knit$set(root.dir = mainDir)

```


## On this page

Biological insights and take-home messages are at the bottom of the page at section Lesson Learnt: @sec-lesson-learnt-genome-composition.

* Here we sequence 37 genomes of Farmhouse yeasts isolates and we control the quality of the sequencing libraries.
* We characterize the phylogenetic origin of the isolates by aligning the reads to six *Saccharomyces* species reference genomes.
* Then, we estimate the ploidy estimations based on their k-mer frequency distributions.

## Origin of the Farmhouse yeasts isolates

Samples were obtained from 44 Northern and Eastern European farmhouse yeast cultures, including from Western & Eastern Norway, Latvia, Lithuania, and Russia. To investigate genetic relationships within and between cultures, we performed whole-genome sequencing on 37 representative isolates from 22 farmhouse yeast cultures of diverse geographical origins.

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 7.5
#| fig.cap: '[Figure 1: Georgaphical distribution of Farmhouse yeasts cultures.]{style="color: steelblue;"}'
panel_a = ggplot2::ggplot() + ggplot2::annotation_custom(
  grid::rasterGrob(
    magick::image_read("data/p02-01/fig-1A-map-with-numbers.png"),
    width = ggplot2::unit(1, "npc"),
    height = ggplot2::unit(1, "npc")),
  -Inf, Inf, -Inf, Inf)

panel_a
```


## Reads QC and Filtering

Libraries QC is performed by trimming low quality bases with Trimmomatic and discard short reads. I keep only properly paired reads and I discard the unpaired, since the latter are just a tiny fraction of my total data, they complicate the analysis and there is not much added value in them (at least for the analysis we run).

The file sample.lst just contains the names of my kveiks samples.


```{bash, eval=FALSE}
# generate FastQC report
for file in *.fastq; do
  ~/bin/FastQC/fastqc $file &
done

# LowQ base trimming and filtering
while read line; do
	java -jar ~/bin/trimmomatic/trimmomatic.jar PE \
		-threads 72 \
		-phred33 \
		$line.R1.fq.gz $line.R2.fq.gz \
		$line.R1.tr.fq.gz $line.R1.tr.un.fq.gz $line.R2.tr.fq.gz $line.R2.tr.un.fq.gz \
		SLIDINGWINDOW:10:30 \
		TRAILING:30 \
		MINLEN:50
done < sample.lst
```



## Genome Composition

We check the composition of the sequenced kveiks by aligning the reads to a set of *Saccharomyces* reference genomes. We add as well as a reference, 7 additional Farmhouse yeasts genomes sequenced by [Preiss et al., 2018](doi: 10.3389/fmicb.2018.02137): *Granvin1*, *Hornidal1*, *Hornidal2*, *Laerdal2*, *Stordal Ebbergarden1*, *Voss1*.


```{r, echo=FALSE}
matrix.d = cbind(
  c("S. cerevisiae", "S. eubayanus", "S. kudriavzevi", "S. mikatae", "S. paradoxus", "S. uvarum"),
  c("S288C_R64-2-1", "GCF_001298625.1", "gca_000167075", "SRX055454", "ASM207914v1", "ASM224264v1"),
  c("Scere", "Seuba", "Skudr", "Smika", "Spara", "Suvar")
)
colnames(matrix.d) = c("Species", "Genome Accession", "Label")

# print table
DT::datatable(
  matrix.d,
  extensions = c("FixedColumns", "FixedHeader"),
  caption = "Table 1: Yeast Reference Genomes.", 
  options = list(
    scrollX = TRUE, paging = TRUE, fixedHeader = FALSE, pageLength = 10
  )
)

```


First, we align the reads to the six _Saccharomyces_ reference genomes. Then we create 10kb windows with ```bedtools``` and we calculate the average coverage for each window.


```{bash, eval=FALSE}
## index the reference genome
~/bin/bwa/bwa index Saccharomyces_RefGen.fa

## reads alignment
while read line ; do

	bwa mem -t 72 \
		-K 100000000 \
		Saccharomyces_RefGen.fa \
	  "${line}".R1.tr.fq.gz \
	  "${line}".R2.tr.fq.gz \
    > "${line}".align.sam
    
  samtools view -@ 72 -Sb \
    "${line}".align.sam > "${line}".align.bam
    
  samtools sort -@ 72 
    "${line}".align.bam "${line}".align.sort
    
done < ../sample.lst 

## bam2bed
# prepare reference genomes
samtools-1.9/samtools faidx Saccharomyces_RefGen.fa
gatk CreateSequenceDictionary --R Saccharomyces_RefGen.fa
cut -f 2,3 Saccharomyces_RefGen.dict \
  | tail -n +2 \
  | sed 's/LN://g' \
  | sed 's/SN://g' \
  > Saccharomyces_RefGen.bedchor
bedtools makewindows \
  -g Saccharomyces_RefGen.bedchor \
	-w 1000 \
	> Saccharomyces_RefGen.1kb_win.tab

# calculate mean coverage
while read line; do
	bedtools coverage \
	  -a Saccharomyces_RefGen.1kb_win.tab \
	  -b "${line}".align.sort.bam -mean \
	  > "${line}".align.sort.1kb_cov.bed;
done

# format output
for file in *.1kb_cov.bed ; do
	sed -i 's/Scere_/Scere\t/g' $file;
	sed -i 's/Seuba_/Seuba\t/g' $file;
	sed -i 's/Skudr_/Skudr\t/g' $file;
	sed -i 's/Smika_/Smika\t/g' $file;
	sed -i 's/Spara_/Spara\t/g' $file;
	sed -i 's/Suvar_/Suvar\t/g' $file;
	python bed_remove0coverage.py \
	  --input $file \
	  | sed '/^[[:space:]]*$/d' \
	  > $(basename $file .bed).no0.bed;
done

for file in *.no0.bed; do
	NAME=$(basename $file .align.sort.1kb_cov.no0.bed);
	sed -i "s/^/$NAME\t/" $file;
done

cat *.1kb_cov.no0.bed > Viking.species.cov.bed


# generate per chromosomes coverages
while read line; do
	OLD=$(echo $line | cut -d " " -f 1)
	NEW=$(echo $line | cut -d " " -f 2)
	sed -i "s/$OLD\t/$NEW\t/g" Viking.species.cov.chr.bed
done < ref_chr_names.lst 

# clean output
for file in *.no0.bed; do
	cp $file $file.1;
	while read line; do
		OLD=$(echo $line | cut -d " " -f 1);
		NEW=$(echo $line | cut -d " " -f 2);
		sed -i "s/$OLD\t/$NEW\t/g" $file.1;
	done < ../00_ref_genomes/ref_chr_names.lst;
	grep Chr $file.1 > $(basename $file .bed).chr.bed;
	rm $file.1 &
done

```


We then remove windows with 0 coverage to have a clearer figure, and we plot the coverages with transforming the y axis in log10 scale, since Krogerus kveiks have much higher coverage than the strains we sequenced.


```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 9
#| fig.cap: '[Figure 2: Coverage of Reference Genomes.]{style="color: steelblue;"}'

data_cov = read.delim("data/p01-01/Viking.species.cov.chr.bed", header = FALSE)
data_cov$V7 = ifelse(data_cov$V6 >= 1, "Up", "Down")

# ylim = 1500
ggplot(data_cov, aes(x = V2, y = V6)) +
  geom_jitter(aes(color = V7),
              show.legend = FALSE,
              alpha = 0.035,
              shape = ".",
              position = position_jitter(0.475)) +
  geom_hline(yintercept = 1, linetype = 2, color = "red") +
  coord_cartesian(expand = FALSE) +
  scale_color_manual(values  = c("grey75", "cyan4")) + 
  scale_y_continuous(limits = c(-4, 9000),
                     trans = "log2",
                     breaks = c(1, 10, 100, 1000),
                     labels = scales::comma) +
  labs(title = "Coverage of Reference Genomes",
       x = "Reference Species",
       y = "Coverage") +
  facet_wrap(~V1, scales = "free_y") +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        panel.background = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "gray75", size = 0.5),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.65))

```



All sequenced farmhouse yeasts are pure _S. cerevisiae_ strains, except for Muri and k7R25 that are _S. cerevisiae_, _S. eubayanus_ and _S. uvarum_ triple hybrids.

We can dig in into the coverage of the single chromosomes and visualize the extent of hybridization and eventual aneuploidies. 



```{r, echo=FALSE, warning=FALSE}

### PER SAMPLE COVERAGE
sample_list = unique(data_cov$V1)
plot_lst = list()

# I iterate through the samples
for (i in 1:length(sample_list)){
  
  # I read the sample name and restrict the plot to that sample
  name = sample_list[i]
  dataset = read.delim(
    paste("data/p01-01/cov/", name, ".align.sort.1kb_cov.no0.chr.bed", sep = ""),
    header = FALSE,
    sep = "\t",
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
  
  # plot
  plot_lst[[name]] = ggplot(dataset, aes(V5, V6, colour = V2)) +
    geom_point(size = 0.2) +
    scale_y_continuous(limits = c(1.5, 9000),
                       trans = "log2",
                       breaks = c(1, 10, 100, 1000),
                       labels = scales::comma) +
    scale_x_discrete(breaks = (seq(0,2e6,10000))) +
    coord_cartesian(expand = FALSE) +
    facet_grid(V2~V3,
               scales = "free",
               space = "free_x",
               switch = "y") +
    labs(title = name,
         x = "Chromosome",
         y = "Genome coverage",
         color = "Genomes") +
    theme(plot.title = element_blank(),
          legend.position = "none",
          panel.background = element_blank(),
          panel.spacing.x = unit(0, "lines"),
          panel.border = element_rect(colour = "black", size = 1, fill = NA),
          strip.background = element_rect(colour = "black", size = 1, fill = NA))
  
}


```


::: panel-tabset

### 14R30

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["14R30"]]

```

### 14R6

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["14R6"]]

```

### 16R23

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["16R23"]]

```

### 16R37

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["16R37"]]

```

### 17P5

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["17P5"]]

```

### 19R18

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["19R18"]]

```

### 1R16

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["1R16"]]

```

### 21P1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["21P1"]]

```

### 21R38

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["21R38"]]

```

### 27R17

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["27R17"]]

```

### 28P1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["28P1"]]

```

### 28P6

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["28P6"]]

```

### 28R21

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["28R21"]]

```

### 28R33

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["28R33"]]

```

### 28R8

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["28R8"]]

```

### 2R23

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["2R23"]]

```

### 38R16

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["38R16"]]

```

### 39R20

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["39R20"]]

```

### 3R11

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["3R11"]]

```

### 40R14

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["40R14"]]

```

### 40R1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["40R1"]]

```

### 40R20

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["40R20"]]

```

### 41R10

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["41R10"]]

```

### 41R15

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["41R15"]]

```

### 42R20

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["42R20"]]

```

### 42R31

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["42R31"]]

```

### 44R32

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["44R32"]]

```

### 44R7

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["44R7"]]

```

### 45P5

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["45P5"]]

```

### 45R11

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["45R11"]]

```

### 46R12

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["46R12"]]

```

### 46R37

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["46R37"]]

```

### 6R15

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["6R15"]]

```

### 7R7

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["7R7"]]

```

### 8R19

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["8R19"]]

```

### 9R40

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["9R40"]]

```

### Granvin1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Granvin1"]]

```

### Hornindal1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Hornindal1"]]

```

### Hornindal2

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Hornindal2"]]

```

### k7R15

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["k7R15"]]

```

### k7R25

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["k7R25"]]

```

### Laerdal2

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Laerdal2"]]

```

### Muri

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Muri"]]

```

### SortdalEbbe1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["SortdalEbbe1"]]

```

### Voss1

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Subgenome coverages.]{style="color: steelblue;"}'

plot_lst[["Voss1"]]

```

:::


*Muri* and *k7R25* are triple hybrids _S. cerevisiae_ X _S. eubayanus_ X _S. uvarum_. What is extremely interesting is that they share the same genomic composition in terms of origin of the chromosomes, but they have distinct CNVs patterns. I would suggest a common origin for such hybrid.



## Ploydies estimation

We can estimates Kveiks ploidy by looking at the k-mer frequency distributions of the sequenced reads obtained from whole genome sequencing.

```{bash, eval=FALSE}

# calculate kmer freq
for file in 01_alignments/*bam; do
  ~/bin/ntCard/bin/ntcard \
    --threads=64 \
    --kmer=41 \
    --pref=$(basename $file .align.sort.md.r.bam).kmer41freq \
    $file;
done

# merge outputs
for file in *.hist; do
  tail -n +3 $file > $file.1;
  mv $file.1 $file;
  cat $file | sed "s/^/$(basename $file .kmer41freq_k41.hist)\t/g";
done > Viking.species.kmer41freq.txt

```


K-mer frequency distribution plots can suggest us the degree of heterozygosity of kveiks genomes. All Kveiks seems to have a high degree of heterozygosity, except for hybrid strains Muri and k7R25. This support the idea that the sequenced kveiks so far are tetraploid _S. cerevisiae_, while Muri and k7R25 are triploid hybrids _S. cerevisiae_ X _Seubayanus_ X _S. uvarum_.


```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 9
#| fig.cap: '[Figure 4: K-mer frequencies distribution.]{style="color: steelblue;"}'

data_k = read.delim("./data/p01-01/Viking.species.kmer41freq.mod.txt", header = FALSE)
data_k$V2 = as.numeric(data_k$V2)
data_k$V3 = as.numeric(data_k$V3)

# plot kmer distributions
ggplot(data_k, aes(x = V2, y = V3)) +
  geom_line(color = "steelblue", size = 1) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  facet_wrap(~V1, scales = "free") +
  labs(title = "K-mer frequencies distribution",
       x = "k-mer depth",
       y = "k-mer frequency") +
  theme(plot.title = element_text(size = 20, hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA),
        strip.background = element_rect(colour = "black", fill = NA),
        strip.text = element_text(size = 8))

```



## Lessons Learnt {#sec-lesson-learnt-genome-composition}

Based on the initial analysis of the Farmhouse yeasts isolates genomics libraries we have learnt:

* All Farmhouse yeasts isolates sequenced so far are pure *S. cerevisiae*, except for *Muri* and the *k7R25* isolate we sequenced. *Muri* was first identified by [Preiss to be a unique *Saccharomyces cerevisiae* × *Saccharomyces eubayanus* x *Saccharomyces uvarum* triple hybrid](https://doi.org/10.3389/fmicb.2018.02253), potentially originating from a contamination of British Ales. *k7R25*, surprisingly, show the same *S. cerevisiae* × *S. eubayanus* x *S. uvarum* hybridization pattern, suggesting a common origin with *Muri*.
* Most farmhouse yeast isolates we analysed appear to be tetraploid based on their k-mer frequency distributions. Exceptions to this are the isolates *16R23*, *38R16*, *39R20* and *40R14* of Lithuanian and Russian origin, which appears to be diploid, and the triple hybrids *Muri* and *k7R25*, which seems to have a single copy of each subgenome.


## Session Information

:::{.callout-note collapse=true}

```{r, echo=FALSE}
utils::sessionInfo()
```

:::
