# High-throughput phenotyping

```{css, echo=FALSE}

pre {
  max-height: 300px;
}

div.sourceCode {
  overflow: scroll;
}

.bash-code {
  font-style: italic;
}


```


```{r, echo=FALSE, include=FALSE}
# load required libraries
up_packages = c("ape", "aplot", "dplyr", "ggnewscale", "ggplot2", "ggpubr", "ggtree", "ggtreeExtra",
                "grid", "gridExtra", "gtable", "mixOmics", "naturalsort", "PCAtools", "phytools", "plyr", "RColorBrewer",
                "reshape", "reshape2", "stringr", "treeio")
lapply(up_packages, require, character.only = TRUE)

# set working directory
mainDir = "/home/andrea/19ADC99/vikingsproject"
knitr::opts_knit$set(root.dir = mainDir)

```


## On this page

Biological insights and take-home messages are at the bottom of the page at section Lesson Learnt: @sec-lesson-learnt-feature-selection.

* Here


## High-throughput phenotyping of farmhouse and industrial yeasts

Let's import the phenotypic data and reformat them.

```{r, warning=FALSE, message=FALSE}

# import and prep data
raw_data = read.delim("./data/p01-06/grouped phenotype data with annotations.csv", sep = ",", header = TRUE)

raw_data = raw_data %>%
  dplyr::filter(group != "WL004") %>%
  dplyr::filter(group != "LA001")
raw_data$Industry = ifelse(
  raw_data$region %in% c("South-West Norway", "Lithuania", "North-West Norway", "Russia",
                         "Central-Eastern Norway", "Latvia", "South-Eastern Norway"),
  "Farmhouse",
  raw_data$region
)
raw_data$Industry = ifelse(
  raw_data$Outside == "yes",
  "Allochthonous\nyeast",
  raw_data$Industry
)
raw_data$region = ifelse(
  raw_data$Industry == "Farmhouse",
  raw_data$region,
  "NA"
)

raw_data$region = ifelse(
  raw_data$culture %in% c("7", "38"),
  "South-West Norway",
  ifelse(
    raw_data$culture == "40",
    "Russia",
    ifelse(
      raw_data$culture == "45",
      "Latvia",
      ifelse(
        raw_data$culture == "57",
        "Central-Eastern Norway",
        raw_data$region
      )
    )
  )
)


raw_data = raw_data %>%
  dplyr::mutate(region = dplyr::case_when(
    region == "South-West Norway" ~ "SW Norway",
    region == "North-West Norway" ~ "NW Norway",
    region == "Central-Eastern Norway" ~ "CE Norway",
    region == "South-Eastern Norway" ~ "SE Norway",
    .default = as.character(region)
  ))

metadata = raw_data %>%
  dplyr::select(c("group", "culture", "region", "Industry", "Outside", "method"))
rownames(metadata) = metadata$group

rownames(raw_data) = raw_data$group
counts = raw_data %>%
  dplyr::select(-c("group", "culture", "region", "Industry", "Outside", "method"))

# import final clade list
final_clades = read.table(
  "./data/p01-06/final_clades_for_pub.txt",
  sep = "\t",
  header = TRUE,
  stringsAsFactors = FALSE
)
# replace
for(i in 1:nrow(final_clades)){

  strain = final_clades[i, "Strain"]
  clade = final_clades[i, "Clade"]
  metadata[which(metadata$group == strain), "Industry"] = clade

}
metadata$Industry = ifelse(metadata$Industry == "Beer", "Beer2", metadata$Industry)
metadata$Industry = ifelse(metadata$group == "BI001", "Asia", metadata$Industry)
metadata$Industry = ifelse(metadata$group == "BR003", "Mixed", metadata$Industry)
metadata$Industry = ifelse(metadata$group == "BI002", "Other", metadata$Industry)
metadata$Industry = ifelse(metadata$group == "BI004", "Asia", metadata$Industry)
metadata$Industry = ifelse(metadata$group == "BI005", "Other", metadata$Industry)


## background: no farmhouse
metadata_no_farm = metadata[which(metadata$Industry != "Farmhouse"), ]
metadata_no_farm = metadata_no_farm[which(metadata_no_farm$Industry != "Allochthonous\nyeast"), ]
counts_no_farm = counts[which(rownames(counts) %in% metadata_no_farm$group), ]

## farmhouse focus
metadata_farm_only = metadata[which(metadata$Industry %in% c("Farmhouse", "Allochthonous\nyeast")), ]
colnames(metadata_farm_only)[5] = "Allochthonous"
metadata_farm_only$Allochthonous = ifelse(metadata_farm_only$Allochthonous == "yes", "yes", "no")
counts_farm_only = counts[which(rownames(counts) %in% metadata_farm_only$group), ]

```


### Principal Component Analysis



#### Background: no farmhouse yeasts


```{r, warning=FALSE, message=FALSE}

# create PCA object
pca_obj = PCAtools::pca(t(counts_no_farm),
                        metadata_no_farm,
                        removeVar = 0.1)


### IDENTIFY NUMBER OF SIGNIFICANT PRINCIPAL COMPONENTS
# Elbow method
elbow_method = PCAtools::findElbowPoint(pca_obj$variance)

# minimum number of components that explains 80% of variance
nPCopt = which(cumsum(pca_obj$variance) > 80)[1]

# max number of component + 1
if(elbow_method >= nPCopt){
  nPCmax = elbow_method + 1
} else{
  nPCmax= nPCopt
}

```

saassa

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 1: Screeplot of PC variants without farmhouse yeasts.]{style="color: steelblue;"}'

# screeplot
p1 = PCAtools::screeplot(pca_obj,
                        components = pca_obj$components[1:nPCmax],
                        vline = c(elbow_method, nPCopt),
                        colBar = rev(colorRampPalette(brewer.pal(7, "Blues"))(nPCmax)),
                        axisLabSize = 18,
                        titleLabSize = 22) +
  geom_label(aes(x = elbow_method, y = 50, label = "Elbow method", vjust = -1, size = 8)) +
  geom_label(aes(x = nPCopt, y = 50, label = "80% variance", vjust = -1, size = 8), colour = "steelblue") +
  theme(plot.title = element_blank())

p1

```


asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 2: Biplot of industrial yests phenotypes.]{style="color: steelblue;"}'

### BIPLOT
p1 = PCAtools::biplot(pca_obj,
                      labSize = 0,
                      pointSize = 5,
                      drawConnectors = FALSE,
                      max.overlaps = 700,
                      colby = "Industry",
                      colkey = c('#e6194b', '#ffe119', '#46f0f0', '#911eb4', "grey75", '#bcf60c'),
                      legendPosition = "right") +
  theme(plot.title = element_blank(),
        plot.subtitle = element_blank(),
        legend.background = element_rect(colour = "grey25", size = 0.5))
p1

```


saasas



```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 3: Biplot of industrial yests phenotypes.]{style="color: steelblue;"}'

## PAIRSPLOT
max_pairplot_comp = ifelse(nPCmax - 1 <= 10, nPCmax - 1, 10)
p1 = PCAtools::pairsplot(pca_obj,
                          components = pca_obj$components[1:(max_pairplot_comp)],
                          colby = "Industry",
                          colkey = c('#e6194b', '#ffe119', '#46f0f0', '#911eb4', "grey75", '#bcf60c'),
                          pointSize = 2,
                          trianglelabSize = 12,
                          plotaxes = FALSE,
                          margingaps = unit(c(0.05, 0.05, 0.05, 0.05), "cm")) +
  theme(plot.title = element_blank())
p1

```



```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 4: Industrial yests phenotypes loadings.]{style="color: steelblue;"}'


### PRINCIPAL COMPONENTS LOADINGS
# plot loadings
p1 = PCAtools::plotloadings(pca_obj,
                            components = getComponents(pca_obj, seq_len(nPCmax - 1)),
                            rangeRetain = 0.01,
                            labSize = 3,
                            title = "Loadings plot: correlation between phenotypes",
                            subtitle = paste0("and the ", nPCmax - 1, " Significant Principal Components "),
                            caption = "Top 1% variables",
                            shape = 21,
                            col = c('steelblue', 'white', 'red3'),
                            drawConnectors = TRUE) +
  theme(plot.title = element_blank(),
        plot.subtitle = element_blank())
p1

```




#### Farmhouse and industrial yeasts


```{r, warning=FALSE, message=FALSE}

# create PCA object
pca_obj = PCAtools::pca(t(counts),
                        metadata,
                        removeVar = 0.1)


### IDENTIFY NUMBER OF SIGNIFICANT PRINCIPAL COMPONENTS

# Elbow method
elbow_method = PCAtools::findElbowPoint(pca_obj$variance)

# minimum number of components that explains 80% of variance
nPCopt = which(cumsum(pca_obj$variance) > 80)[1]

# max number of component + 1
if(elbow_method >= nPCopt){
  nPCmax = elbow_method + 1
} else{
  nPCmax= nPCopt
}

```

sssss

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 5: Screeplot of PC variants of industrial and farmhouse yeasts.]{style="color: steelblue;"}'

# screeplot
p_all_scree = PCAtools::screeplot(pca_obj,
                        components = pca_obj$components[1:nPCmax],
                        vline = c(elbow_method, nPCopt),
                        title = "Screeplot - number of significant Principal Components",
                        ylab = "Explained\nVariation (%)",
                        colBar = rev(colorRampPalette(brewer.pal(7, "Blues"))(nPCmax)),
                        axisLabSize = 18,
                        titleLabSize = 22) +
  geom_label(aes(x = elbow_method, y = 50, label = "Elbow method", vjust = -1, size = 8)) +
  geom_label(aes(x = nPCopt, y = 50, label = "80% variance", vjust = -1, size = 8), colour = "steelblue") +
  theme(plot.title = element_blank(),
        axis.title.x = element_blank())
p_all_scree

```


asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'

### BIPLOT
p_all_bi = PCAtools::biplot(pca_obj,
                            labSize = 0,
                            pointSize = 1.5,
                            drawConnectors = FALSE,
                            max.overlaps = 700,
                            shape ="Outside",
                            shapekey = c(20, 21),
                            colby = "Industry",
                            colkey = c("red", '#e6194b', '#ffe119', '#3cb44b', '#4363d8',  '#e6beff', '#911eb4', '#46f0f0', '#bcf60c'),
                            title = "PCA - Industry",
                            legendPosition = "right") +
  theme(plot.title = element_blank(),
        legend.background = element_rect(colour = "grey25", size = 0.5))
p_all_bi

```

sssssss

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'

## PAIRSPLOT
p_all_pair = PCAtools::pairsplot(pca_obj,
                                 components = pca_obj$components[1:4],
                                 shape ="Outside",
                                 shapekey = c(20, 21),
                                 colby = "Industry",
                                 colkey = c("red", '#e6194b', '#ffe119', '#3cb44b', '#4363d8',  '#e6beff', '#911eb4', '#46f0f0', '#bcf60c'),
                                 pointSize = 1,
                                 trianglelabSize = 12,
                                 plotaxes = FALSE,
                                 margingaps = unit(c(0.05, 0.05, 0.05, 0.05), "cm")) +
  theme(plot.title = element_blank())
print(p_all_pair)

```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'


### PRINCIPAL COMPONENTS LOADINGS
# plot loadings
p_all_loads = PCAtools::plotloadings(pca_obj,
                                     components = getComponents(pca_obj, seq_len(4)),
                                     rangeRetain = 0.01,
                                     labSize = 3,
                                     caption = "Top 1% variables",
                                     shape = 21,
                                     col = c('steelblue', 'white', 'red3'),
                                     drawConnectors = TRUE) +
  theme(plot.title = element_blank(),
        plot.subtitle = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank())
print(p_all_loads)

```



#### Farmhouse yeasts only


```{r, warning=FALSE, message=FALSE}

# create PCA object
pca_obj = PCAtools::pca(t(counts_farm_only),
                        metadata_farm_only,
                        removeVar = 0.1)


### IDENTIFY NUMBER OF SIGNIFICANT PRINCIPAL COMPONENTS

# Elbow method
elbow_method = PCAtools::findElbowPoint(pca_obj$variance)

# minimum number of components that explains 80% of variance
nPCopt = which(cumsum(pca_obj$variance) > 80)[1]

# max number of component + 1
if(elbow_method >= nPCopt){
  nPCmax = elbow_method + 1
} else{
  nPCmax= nPCopt
}

```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'


### SCREEPLOT
# screeplot
p1 = PCAtools::screeplot(pca_obj,
                        components = pca_obj$components[1:nPCmax],
                        vline = c(elbow_method, nPCopt),
                        title = "Screeplot - number of significant Principal Components",
                        colBar = rev(colorRampPalette(brewer.pal(7, "Blues"))(nPCmax)),
                        axisLabSize = 18,
                        titleLabSize = 22) +
  geom_label(aes(x = elbow_method, y = 50, label = "Elbow method", vjust = -1, size = 8)) +
  geom_label(aes(x = nPCopt, y = 50, label = "80% variance", vjust = -1, size = 8), colour = "steelblue") +
  theme(plot.title = element_text(size = 18, hjust = 0.5))
print(p1)


```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'


### BIPLOT
p_farm = PCAtools::biplot(pca_obj,
                          labSize = 0,
                          pointSize = 2,
                          drawConnectors = FALSE,
                          max.overlaps = 700,
                          shape ="Allochthonous",
                          shapekey = c(20, 21),
                          colby = "region",
                          colkey = c('#0571B0', '#FBA01D',"#FFDA00", "steelblue", '#A6611A',"#008470",'#92C5DE'),
                          title = "PCA - region",
                          legendPosition = "right") +
  theme(plot.title = element_blank(),
        plot.subtitle = element_blank(),
        legend.background = element_rect(colour = "grey25", size = 0.5))
plot(p_farm)



```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'



p_farm_pair = PCAtools::pairsplot(pca_obj,
                                  components = pca_obj$components[1:4],
                                  colby = "region",
                                  shape ="Allochthonous",
                                  shapekey = c(20, 21),
                                  colkey = c('#0571B0', '#FBA01D',"#FFDA00", "steelblue", '#A6611A',"#008470",'#92C5DE'),
                                  pointSize = 1,
                                  trianglelabSize = 12,
                                  plotaxes = FALSE,
                                  margingaps = unit(c(0.05, 0.05, 0.05, 0.05), "cm")) +
  theme(plot.title = element_blank())
print(p_farm_pair)

```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'


p1 = PCAtools::biplot(pca_obj,
                      labSize = 0,
                      pointSize = 2,
                      drawConnectors = FALSE,
                      max.overlaps = 700,
                      colby = "method",
                      colkey = c("grey75", "firebrick", "steelblue"),
                      legendPosition = "bottom") +
  theme(plot.title = element_blank(),
        plot.subtitle = element_text(size = 18, hjust = 0.5),
        legend.background = element_rect(colour = "grey25", size = 0.5))
plot(p1)

```

as
as
as
asassaas
as




```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'

p1 = PCAtools::pairsplot(pca_obj,
                          components = pca_obj$components[1:10],
                          colby = "method",
                          colkey = c("grey75", "firebrick", "steelblue"),
                          pointSize = 1,
                          trianglelabSize = 12,
                          plotaxes = FALSE,
                          margingaps = unit(c(0.05, 0.05, 0.05, 0.05), "cm")) +
  theme(plot.title = element_blank())
print(p1)



```

saassasa
asas
asasas

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| fig.height: 6
#| fig.cap: '[Figure 6: Biplot of framhouse and industrial yests phenotypes.]{style="color: steelblue;"}'



### PRINCIPAL COMPONENTS LOADINGS
# plot loadings
p1 = PCAtools::plotloadings(pca_obj,
                            components = getComponents(pca_obj, seq_len(nPCmax - 1)),
                            rangeRetain = 0.01,
                            labSize = 3,
                            caption = "Top 1% variables",
                            shape = 21,
                            col = c('steelblue', 'white', 'red3'),
                            drawConnectors = TRUE)
print(p1)

```




### Discriminant analysis


sPLS-DA


## FArmhouse


### PCA



### Distributions







## Fermentations





## Lessons Learnt {#sec-lesson-learnt-feature-selection}

Based on the  we have learnt:

* Fr


## Session Information

:::{.callout-note collapse=true}

```{r, echo=FALSE}
utils::sessionInfo()
```

:::